Expanding the scope of experimental archaeology using the Perception-Process-Product conceptual framework

### 2023-03-15 manuscript/manuscript.Rmd
added some text.

We believe that contemporary practices in experimental archaeology, as manifested by the fact the the the majority of scholarly publications are produced as results of experiments conducted by single knapper with strong academic background, tend to be restrained by the cognitive bias known as the "curse of knowledge" or "curse of expertise". The curse of knowledge refers to the phenomenon that it is extremely challenging for experts to ignore the information that is held by them but not others, particularly novices [@hinds1999; @camerer1989]. When the knapping expertise is gradually formed through multiple years of observations and trial-and-error learning, an expert knapper develops some specific ways of strategic planning, motor habits, preferences of percussor and raw material types, as well as familiarity of various techniques that become unforgettable. The existence of this cognitive bias is not inherently bad, and these many years of experiences should be appreciated and celebrated by experimental archaeologists. However, what is problematic is that the results of replication experiments conducted by these experienced practioners, often in settings of single knapper, has been constantly framed as grandiose generalization regarding the evolution of technology and cognition that masks a huge range of technological diversity.

@@

>>>>

We believe that contemporary practices in experimental archaeology, as manifested by the fact the the the majority of scholarly publications are produced as results of experiments conducted by single knapper with a dual identity of researcher, tend to be restrained by the cognitive bias known as the "curse of knowledge" or "curse of expertise". The curse of knowledge refers to the phenomenon that it is extremely challenging for experts to ignore the information that is held by them but not others, particularly novices [@hinds1999; @camerer1989]. When the knapping expertise is gradually formed through multiple years of observations and trial-and-error learning, an expert knapper develops some specific ways of strategic planning, motor habits, preferences of percussor and raw material types, as well as familiarity of various techniques that become unforgettable. The existence of this cognitive bias is not inherently bad, and these many years of experiences should be appreciated and celebrated by experimental archaeologists. However, what is problematic is that the results of replication experiments conducted by these experienced practioners, often in settings of single knapper, has been constantly framed as grandiose generalization regarding the evolution of technology and cognition that masks a huge range of technological diversity.

@@

In addition to the difficulty in coordination and logistics, the facilitation of large-scale collaborations is often hindered by the current system of research evaluation, where usually only the first author and the senior (last/correspondent) author of a peer-reviewed journal paper will be acknowledged as proper contribution.
